desc: Infimum of strictly decreasing b_n/n^2
source: USA TST 2020/1
tags: [2019-12, oly, medium, alg, sequence, construction, nice]
author: Carl Schildkraut, Milan Haiman

---

Choose positive integers $b_1$, $b_2$, $\ldots$ satisfying \[1=\frac{b_1}{1^2}>\frac{b_2}{2^2}>\frac{b_3}{3^2}>\frac{b_4}{4^2}>\cdots\]
and let $r$ denote the largest real number satisfying $\frac{b_n}{n^2}\ge r$ for all positive integers $n$. What are the possible values of $r$ across all possible choices of the sequence $(b_n)$?

---

The answer is $0\le r\le1/2$.
\setcounter{claim}0
\begin{claim}
    $r=1/2$ works, and is maximal.
\end{claim}
\begin{proof}
    To achieve $r=1/2$, take $b_n=n(n+1)/2$, from which \[\frac{b_n}{n^2}=\frac{n(n+1)}{2n^2}=\frac{n+1}{2n}=\frac12+\frac1{2n},\]
    which clearly satisfies the problem condition. We inductively show that $b_n\le n(n+1)/2$. The base case has been given to us. Now, if the hypothesis holds for all integers less than $n$, then \[\frac{b_n}{n^2}<\frac{b_{n-1}}{(n-1)^2}\le\frac n{2(n-1)}\implies b_n<\frac{n^3}{2(n-1)}.\]
    It is easy to verify the largest possible $b_n$ is $n(n+1)/2$, as claimed.
\end{proof}
\begin{claim}
    All $r<1/2$ work.
\end{claim}
\begin{proof}
    Consider the sequence $(a_n)$ defined by $a_n:=\left\lceil kn^2\right\rceil+n$. Since $a_n$ is $O(n^2)$ and $k<1/2$, there exists $N$ such that for all $n\ge N$, $a_n/n^2<1/2$. I claim the sequence \[b_n:=\begin{cases}n(n+1)/2&\text{ for }n<N\\ a_n&\text{ for }n\ge N\end{cases}\]
    works. By definition of $N$, $b_n/n^2>b_{n+1}/(n+1)^2$ for $n<N$, so it suffices to verify $a_n/n^2$ is strictly decreasing for $n\ge N$.

    In other words, we want to show that \[L:=\frac{\left\lceil kn^2\right\rceil+n}{n^2}>\frac{\left\lceil k(n+1)^2\right\rceil+n+1}{(n+1)^2}=:R\]
    for all $n\ge N$. Since $\left\lceil kn^2\right\rceil\ge kn^2$, \[L\ge\frac{kn^2+n}{n^2}=k+\frac1n,\]
    and similarly since $\left\lceil k(n+1)^2\right\rceil<k(n+1)^2+1$, \[R<\frac{k(n+1)^2+n+2}{(n+1)^2}=k+\frac{n+2}{(n+1)^2},\]
    so it suffices to verify that \[\frac1n\ge\frac{n+2}{(n+1)^2}\iff(n+1)^2\ge n(n+2),\]
    which is true.
\end{proof}

Combining these two claims, we are done.
