desc: G'' isomorphic to G
source: USA TST 2020/4
tags: [2020-03, oly, hard, combo, graph, anti, trash, invariant]

---

For a finite simple%\footnote{A \emph{finite simple graph} $G=(V,E)$ is a finite set $V$ of vertices, together with a set $E$ of edges, where each edge in $E$ is a set of two distinct vertices of $V$. If $v$ is a vertex of $G$, the \emph{neighbors} of $v$ are the vertices $u$ for which $\{u,v\}\in E$. Two graphs $G_1=(V_1,E_1)$ and $G_2=(V_2,E_2)$ are \emph{isomorphic} if there exists a bijection $f:V_1\to V_2$ such that $\{u,v\}\in E_1$ if and only if $\{f(u),f(v)\}\in E_2$.}
graph $G$, we define $G'$ to be the graph on the same vertex set as $G$, where for any two vertices $u\ne v$, the pair $\{u,v\}$ is an edge of $G'$ if and only if $u$ and $v$ have a common neighbor in $G$. Prove that if $G$ is a finite simple graph which is isomorphic to $(G')'$, then $G$ is also isomorphic to $G'$.

---

Define a sequence of graphs by $G_0=G$ and $G_n=G_{n-1}'$. I will show that for all connected graphs $G$, if $G\simeq G_n$ for some positive integer $n$, then $G\simeq G_1$. To see how this finishes the problem, note that for any graph $G$, if $G\simeq G''$, then the connected components of $G$ must have been permuted. Keep permuting the connected components until they map to themselves.

Henceforth $G$ is connected.
\setcounter{claim}0
\begin{claim}
    If some vertex $v$ in $G$ has degree $\ge3$, then for all triples of neighbors $v_1$, $v_2$, $v_3$ of $v$, all three edges of $\triangle v_1v_2v_3$ are in $G$.
\end{claim}
\begin{proof}
    Let $t(G)$ be the number of triangles in $G$. Assume for contradiction $\triangle v_1v_2v_3$ is not in $G$. Note that for any triangle $\delta$ in any graph $G$, $\delta$ is also in $G'$. Since $\triangle v_1v_2v_3$ is in $G'$, $t(G_n)\ge t(G')>t(G)$, contradiction.
\end{proof}
\begin{claim}
    If some vertex in $G$ has degree $d\ge3$, then $G$ is a clique.
\end{claim}
\begin{proof}
    We proceed inductively on the number of vertices $n$. The base case, $n=4$, is by Claim 1. Now assume the claim holds for all graphs with less than $n$ vertices.

    Let $v$ be the vertex of minimal degree, and assume that $G$ is not a clique. I claim that if we delete $v$, the resulting graph still contains a vertex of degree $\ge3$. Otherwise, $v$ must be connected to a vertex $w$ of degree $3$, and all vertices not neighbors with $v$ have degree $\le2$, so by Claim 1, $v$ is also connected to the other neighbors $w_1$ and $w_2$ of $w$, so its degree is at least $3$. Unless $v$ is connected to all vertices of $G$ (contradicting the assumption $G$ is not a clique), there is a vertex of degree less than $v$, contradiction.

    By the inductive hypothesis, removing $v$ from the graph results in a clique. Applying Claim 1 on all vertices in the clique proves that $v$ is connected to all other vertices, so $G$ is a clique, as claimed.
\end{proof}

Cliques obviously obey the problem statement, so assume all vertices have degree at most $2$. Then $G$ is either a singleton, a long chain, or a polygon; we ignore the singleton, which is easy to settle. Of these, only for a polygon with an odd number of sides is $G''$ connected. In this case, if $G$ is the polygon $V_1V_2V_3\cdots V_{n-1}V_n$ ($n$ odd), then $G'$ is $V_1V_3V_5\cdots V_{n-2}V_1$, so $G\simeq G'$, and we are done.
