desc: Symmedian lemma
source: USAJMO 2015/5
tags: [2019-10, oly, trivial, geo, symmedian]

---

Let $ABCD$ be a cyclic quadrilateral. Prove that there exists a point $X$ on segment $BD$ such that $\angle BAC=\angle XAD$ and $\angle BCA=\angle XCD$ if and only if there exists a point $Y$ on segment $\overline{AC}$ such that $\angle CBD=\angle YBA$ and $\angle CDB=\angle YDA$.

---

\begin{customenv}{First solution, by angle chasing}
    Since the two directions are identical, assume that $X$ exists. It suffices to show that $Y$ exists.
    \begin{center}
        \begin{asy}
            size(8cm);
            defaultpen(fontsize(10pt));

            pen pri=purple;
            pen sec=magenta;
            pen fil=purple+opacity(0.05);

            pair O, A, B, D, K, C, X, Y, Ap, Cp;
            O=(0, 0);
            A=dir(140);
            B=dir(230);
            D=dir(310);
            K=2*circumcenter(B, O, D)-O;
            C=intersectionpoint(K -- (A+0.01*(K-A)), circle(O, 1));
            X=(B+D)/2; Y=(A+C)/2;
            Ap=2*foot(A, O, K)-A;
            Cp=2*foot(C, O, K)-C;

            filldraw(circle(O, 1), fil, pri);
            draw(A -- B -- C -- D -- A, pri);
            draw(A -- Cp, sec); draw(C -- Ap, sec);
            draw(B -- Y -- D, sec);
            draw(A -- C, pri); draw(B -- D, pri);

            dot("$A$", A, A);
            dot("$B$", B, B);
            dot("$C$", C, C);
            dot("$D$", D, D);
            dot("$X$", X, N);
            dot("$Y$", Y, W);
            dot("$A'$", Ap, Ap);
            dot("$C'$", Cp, Cp);
        \end{asy}
    \end{center}

    The key claim is that this condition is equivalent to $ABCD$ being harmonic. Let rays $AX$ and $CX$ intersect $(ABCD)$ again at $C'$ and $A'$, respectively. Since $\widehat{A'D}=2\angle XCD=2\angle BCA=\widehat{BA}$, $\triangle ABD\cong\triangle A'DB$. Similarly, $\triangle CBD\cong\triangle C'DB$, so $X=\overline{AA'}\cap\overline{CC'}$ lies on the perpendicular bisector of $\overline{BC}$; hence, $X$ is the midpoint of $\overline{BD}$, and $ABCD$ is harmonic.

    If $Y$ denotes the midpoint of $\overline{AC}$, since $\overline{BD}$ is a $B$-symmedian of $\triangle ABC$, $\angle CBD=\angle YBA$. Similarly, $\angle CDB=\angle YDA$, so $Y$ exists.
\end{customenv}
\begin{customenv}{Second solution, by Steiner}
    Let $E=\seg{AC}\cap\seg{BD}$. We will show that $ABCD$ is harmonic iff $X$ exists. But \[\left(\frac{AB}{AD}\right)^2=\frac{EB}{ED}\cdot\frac{XB}{XD}=\left(\frac{CB}{CD}\right)^2,\]
    as desired. 
\end{customenv}
