% Input your problem and solution below.
% Three dashes on a newline indicate the breaking points.

---

The sequence $a_1$, $a_2$, $\ldots$ of integers satisfies the conditions:
\begin{enumerate}[label=(\roman*),itemsep=0em]
    \item $1\le a_j\le2015$ for all $j\ge1$,
    \item $k+a_k\ne\ell+a_\ell$ for all $1\le k<\ell$.
\end{enumerate}
Prove that there exist two positive integers $b$ and $N$ for which
\[\left\vert\sum_{j=m+1}^n\big(a_j-b\big)\right\vert\le1007^2\]
for all integers $m$ and $n$ such that $n>m\ge N$.

---

For each positive integer $i$, draw an arrow from $i$ to $i+a_i$ on the number line. This partitions the positive integers into several chains.
Observe that among any 2015 consecutive integers, each chain must contain at least one number. In particular, the number of chains is finite. Say a chain has been introduced if its first element has appeared in the sequence.

I claim we can set
\begin{itemize}[itemsep=0em]
    \item $b$ equal to the number of chains, and
    \item $N$ the first integer for which all chains have already been introduced.
\end{itemize}

Fix $m$ and $n$, and number the chains 1, \ldots, $b$. For each $i$, let $d_k(i)$ be the least positive integer $j$ such that $a_{i+j}$ is in the $k$th chain. Furthermore, let
\[S(i)=d_1(i)+d_2(i)+\cdots+d_b(i).\]
Observe that $a_j-b=S(j)-S(j-1)$, so
\[\sum_{j=m+1}^n\big(a_j-b\big)=S(n)-S(m).\]

Now we can easily establish
\[1+2+\cdots+b\le S(i)\le1+\underbrace{2015+2014+\cdots}_{b-1\text{ terms}}.\]
Hence $|S(n)-S(m)|\le(b-1)(2015-b)\le1007^2$, as needed.
\begin{remark}
    This is really a juggling problem: a ball launched at time $i$ his airtime $a_i$ and returns at $i+a_i$. Here, $b$ is the number of balls and $N$ is the first time when all the balls have been launched.
\end{remark}

