% Input your problem and solution below.
% Three dashes on a newline indicate the breaking points.
% vim: tw=72

---

A \textit{site} is any point $(x, y)$ in the plane such that $x$ and $y$ are both positive integers less than or equal to 20.

Initially, each of the 400 sites is unoccupied. Amy and Ben take turns placing stones with Amy going first. On her turn, Amy places a new red stone on an unoccupied site such that the distance between any two sites occupied by red stones is not equal to $\sqrt5$. On his turn, Ben places a new blue stone on any unoccupied site. (A site occupied by a blue stone is allowed to be at any distance from any other occupied site.) They stop as soon as a player cannot place a stone.

Find the greatest $K$ such that Amy can ensure that she places at least $K$ red stones, no matter how Ben places his blue stones.

---

The answer is $100$. To achieve this, checkerboard-color the grid and let Amy take half of the black squares. Now we show that for each $4\times4$ grid, Amy can place at most four stones if Bob plays optimally. This is clearly sufficient. Consider the following dissection into $4$-cycles:

\begin{center}
    \begin{asy}
        size(4cm); dotfactor *= 2;

        draw((0,0)--(2,1)--(3,3)--(1,2)--cycle,purple);
        draw((1,0)--(3,1)--(2,3)--(0,2)--cycle,green);
        draw((2,0)--(3,2)--(1,3)--(0,1)--cycle,red);
        draw((3,0)--(2,2)--(0,3)--(1,1)--cycle,blue);

        for (real i = 0; i < 4 - 1e-9; i += 1) {
            for (real j = 0; j < 4 - 1e-9; j += 1) {
                dot((i, j));
            }
        }
    \end{asy}
\end{center}
Whenever Amy plays in this $4\times4$ grid, Bob puts a stone in the opposite vertex in the $4$-cycle Amy's stone belongs to. Thus Amy can put at most one stone in each $4$-cycle Amy's stone belongs to. Thus Amy can put at most one stone in each $4$-cycle, so we are done.
