% Input your problem and solution below.
% Three dashes on a newline indicate the breaking points.

---

Alex and Kelvin play a game on a complete graph with $n$ vertices. Kelvin goes first, and the players take turns selecting either a single edge to remove from the graph, or a single vertex to remove from a graph. (Removing a vertex also removes all edges incident to that vertex.) The player who removes the final vertex wins the game. Assuming optimal play, for which positive integers $n$ does Kelvin have a winning strategy?

---

The answer is $n\equiv1,2\pmod3$.

Say a graph $G$ is \emph{winning} if the first player wins, and \emph{losing} otherwise. Say two vertices $u$, $v$ are \emph{symmetric} if the neighbor sets of $u$, $v$ are the same. In particular, $u$, $v$ are not connected by an edge.
\begin{lemma*}
    Let $u$, $v$ be symmetric vertices in $G$, and let $G'$ denote the graph obtained from deleting $u$, $v$ from $G$. Then $G$ is winning if and only if $G'$ is winning.
\end{lemma*}
\begin{proof}
    In what follows, $G-w$ means we remove the vertex $w$ from $G$, and $G-w_1w_2$ means we remove the edge $w_1\sim w_2$ from $G$.

    We downward induct on the number of vertices of $G$. First assume $G'$ is winning. Then for some vertex $w\in G'$, $G'-w$ is a losing graph. Deleting $u$, $v$ from $G-w$ results in a losing graph. Since $u$, $v$ are still symmetric, by inductive hypothesis $G-w$ is a losing graph, so $G$ is winning.

    Conversely assume $G'$ is losing. Say that on his first move, the first player removes either a vertex $w\in G$ or an edge $w_1\sim w_2$.
    \begin{itemize}
        \item Suppose $w\notin\{x,y\}$ and $G'-w$ is winning. Still, $u$, $v$ are symmetric, so $G-w$ is winning by inductive hypothesis.
        \item Suppose $w_1,w_2\notin\{x,y\}$, and $G'-w_1w_2$ is winning. Still, $u$, $v$ are symmetric, so $G-w_1w_2$ is winning by inductive hypothesis.
        \item Suppose (without loss of generality) $w=x$. Then the second player removes $y$ and the resulting graph, $G'$, is losing for the first player.
        \item Suppose (without loss of generality) $w_1=x$. (Thus $w_2\ne y$ since $x$, $y$ are not adjacent.) The second player removes $w_2\sim y$. Subsequently, removing $x$, $y$ from $G-xw_2-yw_2$ results in $G'$, a losing graph. By inductive hypothesis, $G_0$ is losing.
    \end{itemize}
    Hence $G$ is losing as well.
\end{proof}

It is evident $K_1$, $K_2$ are winning and $K_3$ is losing. To prove the problem, we induct on $n$ again, assuming the result for integers less than $n$.
\begin{itemize}
    \item If $n\equiv1\pmod3$, Kelvin removes any vertex, leaving Alex with $K_{n-1}$, which is losing.
    \item If $n\equiv2\pmod3$, Kelvin removes any edge $w_1\sim w_2$. If the resulting graph is $G$, then $G-w_1-w_2=K_{n-2}$ is losing, so by the lemma, $G$ is losing.
    \item If $n\equiv0\pmod3$, we will show Kelvin loses. If Kelvin removes a vertex, Alex obtains $K_{n-1}$, which is winning. If Kelvin removes an edge $w_1\sim w_2$, say the resulting graph is $G$. Then $G-w_1-w_2=K_{n-2}$ is winning, so by the lemma, $G$ is winning.
\end{itemize}
The inductive step is complete.

