% Input your problem and solution below.
% Three dashes on a newline indicate the breaking points.

---

Let $n\ge3$ be an integer. In a country there are $n$ airports and $n$ airlines operating two-way flights. For each airline, there is an odd integer $m\ge3$, and $m$ distinct airports $c_1$, $\ldots$, $c_m$, where the flights offered by the airline are exactly those between the following pairs of airports: $c_1$ and $c_2$; $c_2$ and $c_3$; \ldots; $c_{m-1}$ and $c_m$; $c_m$ and $c_1$.

Prove that there is a closed route consisting of an odd number of flights where no two flights are operated by the same airline.

---

We proceed by strong induction on $n$, with base case trivial. Let the vertex set be $V$. We will try to construct a spanning tree on the $n$ vertices, where each edge is a distinct color. This will not always be possible, but when it is impossible, we will be able to induct down.

Begin by selecting an edge (i.e.\ tree with two vertices) $v_1\sim v_2$. We iteratively add vertices to our tree. Say at some step in the process, we have a tree on the vertices $v_1$, $\ldots$, $v_k$, with $k<n$, and that the $k-1$ edges cover $k-1$ distinct colors. There are three cases from here:
\begin{enumerate}[label=(\roman*)]
    \item We can find another color whose cycle contains some vertex in $\{v_1,\ldots,v_k\}$, but also contains some vertex in $V\setminus\{v_1,\ldots,v_k\}$. Then there is an edge of this color of the form $v_i\sim w$, where $w\notin\{v_1,\ldots,v_k\}$. Add this vertex $w$ to the tree.
    \item We can find another color whose cycle is contained entirely in $\{v_1,\ldots,v_k\}$. Then there are $k$ cycles on the subgraph with vertices $\{v_1,\ldots,v_k\}$, so we win by the inductive hypothesis.
    \item No other colors have cycles that contain an element of $\{v_1,\ldots,v_k\}$. In the subgraph consisting of the remaining $n-k$ vertices, there are $n-k$ cycles, so we win by the inductive hypothesis.
\end{enumerate}
If at any point in the process, one of conditions (ii), (iii) holds, then we are already done. Otherwise, we successfully construct a spanning tree with $n-1$ edges of distinct colors.

Let $x_1\sim x_2\sim\cdots\sim x_m\sim x_1$, with $m$ odd, be the cycle formed by the remaining color, and consider the paths from $x_1$ to $x_2$, from $x_2$ to $x_3$, \ldots, from $x_m$ to $x_1$ in the spanning tree. The sum of their lengths must be even, so one of them, say the path from $x_i$ to $x_{i+1}$ (indices modulo $m$), has even length.

Then union this path with the edge $x_i\sim x_{i+1}$ to construct a cycle of odd size all of whose edges are distinct colors. End proof.

