% Input your problem and solution below.
% Three dashes on a newline indicate the breaking points.

---

A configuration of 4027 points in the plane is called Colombian if it consists of 2013 red points and 2014 blue points, and no three of the points of the configuration are collinear. By drawing some lines, the plane is divided into several regions. An arrangement of lines is good for a Colombian configuration if the following two conditions are satisfied:
\begin{enumerate}[label=(\roman*),itemsep=0em]
    \item No line passes through any point of the configuration.
    \item No region contains points of both colors.
\end{enumerate}
Find the least value of $k$ such that for any Colombian configuration of 4027 points, there is a good arrangement of $k$ lines.

---

The answer is 2013.

\bigskip

\textbf{Proof of lower bound:} It suffices to find an arrangement for which 2013 lines are necessary. Just place the 4027 points on the circumference of a circle, so that they alternate colors (with a single exception).

Then the lines must divide the circle into 4026 arcs, i.e.\ 2013 lines are needed.

\bigskip

\textbf{Proof of upper bound:} The key is that we can take any two points and isolate them from the rest by drawing two parallel lines arbitrarily close to them.
\begin{itemize}
    \item If the convex hull contains a red point, isolate it with a line, and then pair up the remaining 2012 red points and isolate them, for a total of $1+2\cdot1006=2013$ lines.
    \item If the convex hull is entirely blue, isolate two adjacent blue points with a line, then pair up the remaining 2012 blue points and isolate them, for a total of $1+2\cdot1006=2013$ lines again.
\end{itemize}
Thus 2013 lines suffice.

