desc: Cars
source: TSTST 2019/3
tags: [2019-10, oly, tricky, combo, troll, beluhov]

---

On an infinite square grid we place finitely many \emph{cars}, which each occupy a single cell and face in one of the four cardinal directions. Cars may never occupy the same cell. It is given that the cell immediately in front of  each car is empty, and moreover no two cars face towards each other (no right-facing car is to the left of a left-facing car within a row, etc.). In a \emph{move}, one chooses a car and shifts it one cell forward to a vacant cell. Prove that there exists an infinite sequence of valid moves using each car infinitely many times.

---

Call a car \emph{horizontal} if it is left-facing or right-facing, and \emph{vertical} otherwise. Color the columns black and white in an alternating manner.

Surround all the cars in a box. It is sufficient to show that the cars can all leave the box. Perform the following procedure:
\begin{itemize}
    \item Move all horizontal cars in a black column to a white column. This is possible because the cell immediately in front of each car is empty.
    \item Clear the black columns of vertical cars. This is possible because all cars in a black column must face the same direction.
    \item Move all the horizontal cars into black columns. This is possible because all vertical cars in black columns have left.
    \item Clear the white columns of vertical cars, which is possible for similar reasons as above.
    \item Clear the box of horizontal cars, which must all face the same direction
\end{itemize}
With the box sufficiently big, all cars can move infinitely many times, as desired.
